---
title: "Carbon Pricing and Decoupling in Europe"
author: "Inhwan Ko, Taedong Lee"
date: 'September 15, 2021'
output: html_document
editor_options: 
  chunk_output_type: console
---

## Ko, Inhwan, and Taedong Lee. Carbon pricing and decoupling between greenhouse gas emissions and economic growth: A panel study of 29 European countries, 1996-2014. 

*Accepted in the Review of Policy Research*

**Inhwan Ko**  
PhD student, Department of Political Science, University of Washington, Seattle, WA 98105

**Taedong Lee, PhD**  
Associate Professor, Political Science Department, Seoul, Korea  

**Abstract**  

This study explores why the levels of decoupling between greenhouse gas (GHG) emissions and economic growth vary across time and countries by examining to which extent carbon pricing instruments which factors are driving this decoupling. We expect that the implementation of carbon pricing policies instruments facilitates decoupling, as they are designed to achieve cost-efficient GHG reduction. We analyze a panel data of 29 European countries between 1996 and 2014 to examine the relationships between two carbon pricing policies instruments (emission trading and carbon tax) and emission intensity (GHG emissions per unit of GDP) which we use to measure decoupling trends. Results from two-way fixed effects models show that emission trading contributes to decoupling, whereas our evidence does not support the role of carbon tax. Furthermore, emission trading is negatively associated with both emission intensity and GHG emissions, implying that it contributes to strong decoupling. Using coarsened exact matched (CEM) data, our results suggests that even a single emission trading policy (e.g., EU-ETS) across different jurisdictions may render a heterogeneous effect on decoupling depending on their socioeconomic conditions. 

**Data**
https://github.com/inhwanko/Carbon-Pricing-and-Decoupling/blob/master/decoupling.csv 

# 1. Load neccessary packages and data

```{r, warning=F, message=F}
library(foreign)
library(tidyverse)
library(plm)
library(cem)
library(ggplot2)
library(ggpubr)
library(estimatr)
library(broom)
library(stargazer)
library(Hmisc)
library(MASS)
library(lmtest)
library(simcf)
library(tile)

# setwd("~")
rawdata <- read.csv("decoupling.csv") ## raw dataset is uploaded on the github 
rawdata <- as_tibble(rawdata)
rawdata <- rawdata[,-1]
rawdata
```

# 2. Data cleaning

```{r, warning=F, message=F}

rawdata$gdpgrowth <- as.numeric(rawdata$gdpgrowth)
data <- rawdata %>% 
  filter(year!=2015 & year!=2016 & year!=2017 & year!=2018)

data <- pdata.frame(data, index=c("countrycode","year"))

data$time <- as.numeric(data$year)-1995
unique(data$countrycode)
```

# 3. Model results

### 3.1. Setting up the formuli

```{r}
colnames(data)
formula1 <- decoup ~ ets + carbontax + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + gdpgrowth + indpergdp
formula2 <- decoup ~ ets*goveff + carbontax*goveff + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + goveff + gdpgrowth + indpergdp

formula3 <- emission ~ ets + carbontax + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + gdpgrowth + indpergdp
formula4 <- emission ~ ets*goveff + carbontax*goveff + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + goveff + gdpgrowth + indpergdp

```

### 3.2. A Hausman test

```{r}

model1f <- plm(formula1, data=data, model="within", effect = "twoway")
model1r <- plm(formula1, data=data, model="random", effect = "twoway")
phtest(model1f, model1r) 

model2f <- plm(formula2, data=data, model="within", effect = "twoway")
model2r <- plm(formula2, data=data, model="random", effect = "twoway")
phtest(model2f, model2r) 

model3f <- plm(formula3, data=data, model="within", effect = "twoway")
model3r <- plm(formula3, data=data, model="random", effect = "twoway")
phtest(model3f, model3r) 

model4f <- plm(formula4, data=data, model="within", effect = "twoway")
model4r <- plm(formula4, data=data, model="random", effect = "twoway")
phtest(model4f, model4r) 

```

### 3.3. Model results

```{r}
summary(model1f, vcov=vcovSCC(model1f, type="HC4", cluster="group"))
summary(model2f, vcov=vcovSCC(model2f, type="HC4", cluster="group"))
summary(model3f, vcov=vcovSCC(model3f, type="HC4", cluster="group"))
summary(model4f, vcov=vcovSCC(model4f, type="HC4", cluster="group"))

```

### 3.4. Interaction for Model 1

```{r}
formula1.2 <- decoup ~ ets + carbontax + co2inten*ets + epdtloss*ets + consumere*ets + consumeff*ets + urbanpopper*ets + polity*ets + gdpgrowth*ets + indpergdp*ets

model1.2f <- plm(formula1.2, data=data, model="within", effect="twoway")
summary(model1.2f, vcov=vcovSCC(model1.2f, type="HC4", cluster="group"))
```



# 4. Coarsened Exact Matching

```{r, warning=F, message=F}
require(cem)

## assigning treatment value (1- ETS, 0- No ETS)

ets1 <- which(data$ets==1)
ets0 <- which(data$ets==0)
nets1 <- length(ets1)
nets0 <- length(ets0)

## checking naive mean difference

mean(data$decoup[ets1])-mean(data$decoup[ets0])

## assigning control variables (including carbon tax)

vars_ets <- c("carbontax", "co2inten","epdtloss","consumere","consumeff","urbanpopper","polity","gdpgrowth","indpergdp")

## you will only need variables to do the matching, without countrycode and year
colnames(data)
data_ets <- data[,c(6,11,12,13,14,15,16,18,20,19,21)]

## deriving univariate imbalance measures:

imbalance(group=data_ets$ets, data=data_ets[vars_ets])

## do matching
mat <- cem(treatment="ets", data=data_ets, drop="decoup")
mat
```

```{r}
ets_linear <- att(mat, decoup~ets, data=data_ets, model="linear")

ets_linear_control <- att(mat, decoup~ets + carbontax + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + gdpgrowth + indpergdp, data=data_ets,
                  model="linear") 

ets_linearRE_control <- att(mat, decoup~ets + carbontax + co2inten + epdtloss + consumere + consumeff + urbanpopper + polity + gdpgrowth + indpergdp, data=data_ets,
                  model="linear-RE") 

ets_linear ## first row of table 
ets_linear_control ## second row of the table
ets_linearRE_control ## third row of the table


```

# 5. Plots

## Figure 1

```{r}
plot1 <- ggplot(data, aes(x=year)) + 
  geom_point(aes(y=emission1996)) +
  geom_line(aes(y=gdp1996, group=1)) +
  labs(x="Year", y="Real GDP of USD 2010 constant and CO2eq emission (1=1996)") +
  theme(axis.title.x=element_text(size=14)) +
  theme(axis.title.y=element_text(size=14)) +
  scale_x_discrete(breaks=c(2000, 2005, 2010)) +
  facet_wrap(~countrycode, nrow=5)
# + ggtitle("Figure 1. Small multiples of decoupling trend in each country, 29 #European countries, 1996-2014")

plot1 + theme_gray(base_size=15)
```

## Figure 2

```{r}
summary(model2f, vcov=vcovSCC(model2f, type="HC4", cluster="group"))

set.seed(2020)
simbeta <- mvrnorm(10000, 
                   coefficients(model2f), 
                   vcovSCC(model2f, type="HC4", cluster="group"))

simbeta <- simbeta[, c(1,2,12)]
simcoef <- colMeans(simbeta)
simlwr <- c(quantile(simbeta[,1],0.025), 
            quantile(simbeta[,2],0.025), 
            quantile(simbeta[,3],0.025))
simupr <- c(quantile(simbeta[,1],0.975), 
            quantile(simbeta[,2],0.975), 
            quantile(simbeta[,3],0.975))

x2 <- seq(-2.5, 2.5, by=0.1)
x1 <- rep(1,length(x2))
y <- simcoef[1]*x1 + simcoef[2]*x2 + simcoef[3]*x1*x2
lwr <- simlwr[1]*x1 + simlwr[2]*x2 + simlwr[3]*x1*x2
upr <- simupr[1]*x1 + simupr[2]*x2 + simupr[3]*x1*x2

interact <- data.frame(y=y, x1=x1, x2=x2, lwr=lwr, upr=upr)

col <- RColorBrewer::brewer.pal(3, "Set3")

figure2.1 <- ggplot(interact, aes(x=x2,y=y)) +
  geom_line(col=col[3], size=1) +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=col[3]), alpha=0.2) +
  theme_minimal(base_size=15) +
  xlab("Government effectiveness score") + 
  ylab("Marginal effect on emission intensity") +
  annotate("text", x=0, y=0, 
           label="When emission trading is in effect", fontface=2, size=5)

figure2.1 <- figure2.1 + theme(legend.position="none")

x2 <- seq(-2.5, 2.5, by=0.1)
x1 <- rep(0,length(x2))
y <- simcoef[1]*x1 + simcoef[2]*x2 + simcoef[3]*x1*x2
lwr <- simlwr[1]*x1 + simlwr[2]*x2 + simlwr[3]*x1*x2
upr <- simupr[1]*x1 + simupr[2]*x2 + simupr[3]*x1*x2

interact <- data.frame(y=y, x1=x1, x2=x2, lwr=lwr, upr=upr)

figure2.2 <- ggplot(interact, aes(x=x2,y=y)) +
  geom_line(col=col[3], size=1) +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=col[3]), alpha=0.2) +
  theme_minimal(base_size=15) +
  xlab("Government effectiveness score") + 
  ylab("Marginal effect on emission intensity") +
  annotate("text", x=0, y=500, 
           label="When emission trading is not in effect", fontface=2, size=5)

figure2.2 <- figure2.2 + theme(legend.position="none")

figure2 <- ggarrange(figure2.1, figure2.2)
figure2
```

## Figure 3
```{r}
plot(ets_linear_control, mat, data_ets, vars_ets)
plot(ets_linearRE_control, mat, data_ets, vars_ets)
```

